### 2024-12-16 NLP/exercises.tex
Updated NLP 2021 exam.

\begin{exercise}{Comprehensiveness}
  Explain comprehensiveness.

  \begin{solution}
    Comprehensiveness refers to the extent to which a system, dataset, or model covers all relevant aspects of a task or domain. In NLP:
    \begin{itemize}
        \item \textbf{Datasets}: Comprehensive datasets should include diverse and representative samples of the target domain.
        \item \textbf{Models}: Comprehensive models should handle all variations and exceptions within their scope.
        \item \textbf{Applications}: Comprehensive applications should meet all user requirements for a given task.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Complete the picture}
  Fill in the white boxes in the image.

  \begin{solution}
    To complete the image, analyze the given context and fill in the white boxes based on:
    \begin{itemize}
        \item \textbf{Labels}: Assign appropriate titles or labels to the boxes (e.g., steps in a pipeline).
        \item \textbf{Relationships}: Connect missing elements to ensure the flow aligns with the diagram's logic.
        \item \textbf{Annotations}: Use any provided hints to accurately reconstruct the intended content.
    \end{itemize}
    The exact answer depends on the image provided.
  \end{solution}
\end{exercise}

\begin{exercise}{NLTK}
  Explain two parameters in NLTK that start with a "c."

  \begin{solution}
    Two common NLTK parameters starting with "c" are:
    \begin{itemize}
        \item \textbf{Chunking}: Groups words into meaningful units (e.g., noun phrases) based on part-of-speech tagging.
        \item \textbf{Concordance}: Displays occurrences of a word within a context window, useful for lexical analysis.
    \end{itemize}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    For multi-class classification, SVM uses:
    \begin{itemize}
        \item \textbf{One-vs-One (OvO)}: Trains a classifier for every pair of classes. Final classification is determined by majority voting.
        \item \textbf{One-vs-All (OvA)}: Trains one classifier per class against all other classes. The highest-confidence prediction is chosen.
        \item \textbf{Extensions}: Kernel tricks and soft margins can be applied for better generalization.
    \end{itemize}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{itemize}
        \item \textbf{Classical ML}:
        \begin{itemize}
            \item Relies on manually engineered features (e.g., TF-IDF, n-grams).
            \item Requires domain expertise for feature extraction.
        \end{itemize}
        \item \textbf{Deep learning}:
        \begin{itemize}
            \item Automatically learns hierarchical feature representations.
            \item Captures complex patterns (e.g., semantics, dependencies).
            \item Requires large datasets for training.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Cohen's Kappa}
  Calculate Cohen's Kappa for a given table.

  \begin{solution}
    Cohen's Kappa is calculated as:
    \[
    \kappa = \frac{P_o - P_e}{1 - P_e}
    \]
    Where:
    \begin{itemize}
        \item \(P_o\): Observed agreement (sum of diagonal elements divided by the total number of items).
        \item \(P_e\): Expected agreement (sum of product of marginal probabilities for each category).
    \end{itemize}
    Substitute the values from the given table into the formula to compute \(\kappa\).
  \end{solution}
\end{exercise}

\begin{exercise}{Evaluation metrics}
  Calculate Precision, Recall, and F1-Score.

  \begin{solution}
    Given true positives (\(TP\)), false positives (\(FP\)), and false negatives (\(FN\)):
    \begin{itemize}
        \item \textbf{Precision}: \( \frac{TP}{TP + FP} \)
        \item \textbf{Recall}: \( \frac{TP}{TP + FN} \)
        \item \textbf{F1-Score}: \( 2 \times \frac{\text{Precision} \times \text{Recall}}{\text{Precision} + \text{Recall}} \)
    \end{itemize}
    Plug in the values provided to compute these metrics.
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    Classification is used in IR for:
    \begin{itemize}
        \item \textbf{Query classification}: Identify user intent (e.g., navigational vs. informational).
        \item \textbf{Document classification}: Categorize documents into predefined topics or genres.
        \item \textbf{Relevance classification}: Predict whether a document is relevant to a query.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Translate datasets}
  List 3 reasons for and against translating an existing annotated dataset from German to another language.

  \begin{solution}
    \begin{itemize}
        \item \textbf{For}:
        \begin{itemize}
            \item Expands dataset availability to resource-scarce languages.
            \item Enables cross-lingual NLP research.
            \item Saves effort in recreating annotations.
        \end{itemize}
        \item \textbf{Against}:
        \begin{itemize}
            \item Potential loss of annotation quality in translation.
            \item Context or cultural relevance may not transfer.
            \item Requires manual review to ensure correctness.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Web corpus}
  List 3 reasons for and against using the web as a corpus.

  \begin{solution}
    \begin{itemize}
        \item \textbf{For}:
        \begin{itemize}
            \item Large-scale, diverse data sources.
            \item Includes up-to-date content.
            \item Covers multiple languages and domains.
        \end{itemize}
        \item \textbf{Against}:
        \begin{itemize}
            \item Noisy and unstructured data.
            \item Contains duplicates and redundant information.
            \item May include biased or low-quality content.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Pipeline}
  Given the pipeline from the first slide, assign three words to the individual steps.

  \begin{solution}
    Example:
    \begin{itemize}
        \item \textbf{Preprocessing}: Tokenization, normalization, and cleaning.
        \item \textbf{Feature extraction}: TF-IDF, embeddings.
        \item \textbf{Modeling}: Training, evaluation, and deployment.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{MRR and accuracy}
  Calculate MRR and accuracy for a Question/Answering system.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Mean Reciprocal Rank (MRR)}:
        \[
        MRR = \frac{1}{N} \sum_{i=1}^{N} \frac{1}{\text{rank}_i}
        \]
        Where \(\text{rank}_i\) is the rank of the first correct answer for query \(i\).
        \item \textbf{Accuracy}: \(\frac{\text{Number of correct answers}}{\text{Total number of queries}}\)
    \end{itemize}
    Calculate based on the provided data.
  \end{solution}
\end{exercise}

\begin{exercise}{Language detection}
  Explain language detection.

  \begin{solution}
    Language detection involves identifying the language of a given text. Common approaches:
    \begin{itemize}
        \item \textbf{N-gram analysis}: Compare character n-grams to language profiles.
        \item \textbf{Lexicon-based methods}: Match text to language-specific dictionaries.
        \item \textbf{Machine learning}: Train classifiers (e.g., SVM, neural networks) on labeled text.
        \item \textbf{Hybrid models}: Combine multiple techniques for higher accuracy.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Plagiarism detection}
  Explain how to find plagiarism.

  \begin{solution}
    Plagiarism detection methods:
    \begin{itemize}
        \item \textbf{String matching}: Use algorithms to find exact matches between documents.
        \item \textbf{Semantic similarity}: Apply embeddings to detect paraphrased or rephrased content.
        \item \textbf{Citation analysis}: Check for missing or incorrect citations.
        \item \textbf{Fingerprinting}: Generate hashes of text fragments and compare across sources.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Crowdsourcing}
  List the pros and cons of crowdsourcing.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Pros}:
        \begin{itemize}
            \item Fast and scalable data annotation.
            \item Access to diverse annotators.
            \item Cost-effective compared to hiring experts.
        \end{itemize}
        \item \textbf{Cons}:
        \begin{itemize}
            \item Quality control is challenging.
            \item May introduce noise or bias.
            \item Difficult to ensure annotator expertise.
        \end{itemize}
    \end{itemize}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    Recall can be improved by:
    \begin{itemize}
        \item \textbf{Expanding queries}: Use synonyms, related terms, or stemming to match more results.
        \item \textbf{Reducing filters}: Remove restrictive criteria (e.g., date, location).
        \item \textbf{Boosting document recall}: Index more documents or reduce relevance thresholds.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Spell checker}
  Explain how a spell checker works with non-words.

  \begin{solution}
    For non-words, spell checkers:
    \begin{itemize}
        \item \textbf{Dictionary lookup}: Identify words not in the dictionary.
        \item \textbf{Edit distance}: Suggest corrections based on similarity (e.g., Levenshtein distance).
        \item \textbf{Contextual models}: Use surrounding words to predict the intended word.
    \end{itemize}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    The Vector Space Model represents documents and queries as vectors:
    \begin{itemize}
        \item \textbf{Terms}: Each dimension corresponds to a term in the vocabulary.
        \item \textbf{Weights}: TF-IDF scores are used as weights for each term.
        \item \textbf{Similarity}: Compute cosine similarity to rank documents by relevance to the query.
    \end{itemize}
  \end{solution}
\end{exercise}

@@

\begin{exercise}{Sentiment analysis}
  Explain the areas where sentiment analysis is used.

  \begin{solution}
    Sentiment analysis applications:
    \begin{itemize}
        \item \textbf{Customer feedback}: Analyze reviews for product improvements.
        \item \textbf{Social media monitoring}: Detect trends and public sentiment.
        \item \textbf{Political analysis}: Gauge public opinion on policies or events.
        \item \textbf{Market research}: Understand consumer preferences.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Teacher/Student}
  Explain Teacher/Student and why it is better than just audio.

  \begin{solution}
    The Teacher/Student approach combines:
    \begin{itemize}
        \item \textbf{Teacher signals}: Provide detailed instructions and explanations.
        \item \textbf{Student feedback}: Allows iterative learning and personalization.
        \item \textbf{Better than audio}: Interaction and adaptivity enhance understanding and retention compared to static audio content.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Representation and fusion}
  Explain representation and fusion and the differences.

>>>>

\begin{exercise}{Comprehensiveness}
  What is comprehensiveness, specifically in the context of NLP?

  \begin{solution}
    Comprehensiveness refers to the extent to which a system, dataset, or model covers all relevant aspects of a task or domain.
    In NLP, comprehensiveness can be evaluated based on the coverage of linguistic phenomena, language varieties, or domain-specific terms in a dataset or model.
  \end{solution}
\end{exercise}

\begin{exercise}{Complete the picture}
  \input{res/qa.fig}

  Fill in the five white boxes in the image.

  \begin{solution}
    1. query formulation, 2. query classification, 3. indexing, 4. doc retrieval, 5. passage retrieval
  \end{solution}
\end{exercise}

\begin{exercise}{NLTK}
  Explain two parameters in NLTK that start with a "c".

  \begin{solution}
    \begin{itemize}
      \item \textbf{Chunking}: Identifies and groups words into syntactic chunks based on their POS tags.
      \item \textbf{Collocations}: Identifies words that frequently co-occur within a certain span.
      \item \textbf{Concordance}: Displays a word in context, showing its surrounding words within a specified window.
    \end{itemize}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{itemize}
        \item \textbf{One-vs-One (OvO)}: Trains a classifier for every pair of classes. Final classification is determined by majority voting.
        \item \textbf{One-vs-All (OvA)}: Trains one classifier per class against all other classes. The highest-confidence prediction is chosen.
    \end{itemize}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{itemize}
        \item \textbf{Classical ML} relies on manually engineered features (like tf-idf or n-grams) and requires domain expertise for feature extraction.
        \item \textbf{Deep networks} automatically learn hierarchical feature representations from raw data, capturing complex patterns and dependencies, but requiring large datasets for training.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Cohen's Kappa}
  \input{res/results.tbl}
  Calculate Cohen's Kappa for Table \ref{tbl:results}.

  \begin{solution}
    Cohen's Kappa is calculated as $\kappa = \frac{P_o - P_e}{1 - P_e}$ and describes the observed agreement $P_o$ relative to the expected agreement $P_e$.
      \[ \kappa = \frac{.7333-.5}{1-.5} = \frac{0.2333}{0.5} = 0.4666 \]
  \end{solution}
\end{exercise}

\begin{exercise}{Evaluation metrics}
  \input{res/results.tbl}
  Calculate Precision, Recall, and F1-Score for Table \ref{tbl:results}.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Precision} $\frac{TP}{TP + FP} = \frac{10}{10+5} = .6666$
        \item \textbf{Recall}: $\frac{TP}{TP + FN} = \frac{10}{10+3} = .7692$
        \item \textbf{F1-Score}: $2 \times \frac{\text{Precision} \times \text{Recall}}{\text{Precision} + \text{Recall}} = 2 \times \frac{.6666 \times .7692}{.6666 + .7692} = .7142$
    \end{itemize}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{itemize}
        \item \textbf{Query classification}: Identify user intent (e.g., navigational vs. informational).
        \item \textbf{Document classification}: Categorize documents into predefined topics or genres.
        \item \textbf{Relevance classification}: Predict whether a document is relevant to a query.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Translate datasets}
  What are the pros and cons of translating datasets for NLP tasks? List 3 for each.

  \begin{solution}
    \begin{itemize}
        \item[+] Expands dataset availability to resource-scarce languages.
        \item[+] Enables cross-lingual NLP research.
        \item[+] Saves effort in recreating annotations.
        \item[-] Potential loss of annotation quality in translation.
        \item[-] Context or cultural relevance may not transfer.
        \item[-] Requires manual review to ensure correctness.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Web corpus}
  What are opportunities and challenges in using the web as a corpus?

  \begin{solution}
    \begin{itemize}
        \item[+] Large-scale, diverse data sources.
        \item[+] Includes up-to-date content.
        \item[+] Covers multiple languages and domains.
        \item[-] Noisy and unstructured data.
        \item[-] Contains duplicates and redundant information.
        \item[-] May include biased or low-quality content.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Pipeline}
  \input{res/pipeline.fig}
  Assign three words to each stage of the given NLP pipeline in Figure \ref{fig:pipeline}.

  \begin{solution}
    \begin{itemize}
      \item \textbf{Data acquisition}: Crawling, scraping, downloading.
      \item \textbf{Text cleaning}: Tokenization, normalization, filtering.
      \item \textbf{Preprocessing}: Lemmatization, POS tagging, parsing.
      \item \textbf{Feature engineering}: TF-IDF, embeddings, n-grams.
      \item \textbf{Modeling}: Training, evaluation, and deployment.
      \item \textbf{Evaluation}: Metrics, analysis, feedback.
      \item \textbf{Deployment}: Integration, monitoring, maintenance.
      \item \textbf{Monitoring \& Model Updating}: Performance tracking, retraining, versioning.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{MRR and accuracy}
  \input{res/qa-answers.fig}
  Calculate MRR and Accuracy for the Question/Answering system in Figure \ref{fig:qa-answers}.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Mean Reciprocal Rank} $MRR = \frac{1}{N} \sum_{i=1}^{N} \frac{1}{\text{rank}_i} = \frac{1}{3} (\frac{1}{3}+\frac{1}{1}+0 = .4444)$
        \item \textbf{Accuracy} $\frac{\text{Number of correct answers}}{\text{Total number of queries}} = \frac{2}{3} = .6666$
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Language detection}
  Explain language detection and common approaches.

  \begin{solution}
    Language detection involves identifying the language of a given text.
    \begin{itemize}
        \item \textbf{N-gram analysis}: Compare character n-grams to language profiles.
        \item \textbf{Lexicon-based methods}: Match text to language-specific dictionaries.
        \item \textbf{Machine learning}: Train classifiers (e.g., SVM, neural networks) on labeled text.
        \item \textbf{Hybrid models}: Combine multiple techniques for higher accuracy.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Plagiarism detection}
  Explain three methods to detect plagiarism in text.

  \begin{solution}
    \begin{itemize}
        \item \textbf{String matching}: Use algorithms to find exact matches between documents.
        \item \textbf{Semantic similarity}: Apply embeddings to detect paraphrased or rephrased content.
        \item \textbf{Citation analysis}: Check for missing or incorrect citations.
        \item \textbf{Fingerprinting}: Generate hashes of text fragments and compare across sources.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Crowdsourcing}
  What are opportunities and challenges in using crowdsourcing for data annotation?

  \begin{solution}
    \begin{itemize}
        \item[+] Fast and scalable data annotation.
        \item[+] Access to diverse annotators.
        \item[+] Cost-effective compared to hiring experts.
        \item[-] Quality control is challenging.
        \item[-] May introduce noise or bias.
        \item[-] Difficult to ensure annotator expertise.
    \end{itemize}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{itemize}
        \item \textbf{Expanding queries}: Use synonyms, related terms, or stemming to match more results.
        \item \textbf{Reducing filters}: Remove restrictive criteria (e.g., date, location).
        \item \textbf{Boosting document recall}: Index more documents or reduce relevance thresholds.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Spell checker}
  How does a spell checker handle non-words?

  \begin{solution}
    \begin{itemize}
        \item \textbf{Dictionary lookup}: Identify words not in the dictionary.
        \item \textbf{Edit distance}: Suggest corrections based on similarity (e.g., Levenshtein distance).
        \item \textbf{Contextual models}: Use surrounding words to predict the intended word.
    \end{itemize}
  \end{solution}
\end{exercise}

@@

  \begin{solution}
    \begin{itemize}
        \item \textbf{Terms}: Each dimension corresponds to a term in the vocabulary.
        \item \textbf{Weights}: TF-IDF scores are used as weights for each term.
        \item \textbf{Similarity}: Compute cosine similarity to rank documents by relevance to the query.
    \end{itemize}
  \end{solution}
\end{exercise}

@@

\begin{exercise}{Sentiment analysis}
  Explain three areas where sentiment analysis is used.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Customer feedback}: Analyze reviews for product improvements.
        \item \textbf{Social media monitoring}: Detect trends and public sentiment.
        \item \textbf{Political analysis}: Gauge public opinion on policies or events.
        \item \textbf{Market research}: Understand consumer preferences.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Teacher/Student}
  Explain the teacher/student signals approach in educational NLP and why it is better than pure audio content.

  \begin{solution}
    \begin{itemize}
        \item \textbf{Teacher signals}: Provide detailed instructions and explanations.
        \item \textbf{Student feedback}: Allows iterative learning and personalization.
        \item \textbf{Better than audio}: Interaction and adaptivity enhance understanding and retention compared to static audio content.
    \end{itemize}
  \end{solution}
\end{exercise}

\begin{exercise}{Representation and fusion}
  What are representation and fusion in multimodal NLP, and how do they differ?
